Final Paper

### 2016-08-08 paper/paper.Rmd
finished paper

Whenever a person hears about a business or think about starting their own, they are always told that a business’s success is all about location, location, location. It’s hard to quantify how much of a business’s success is determined by its location, but we were curious to see if location was linked to higher star ratings on Yelp. Yelp is a service that publishes crowd-sourced reviews about local business online. Additionally, we wanted to see if there was any relationship between certain attributes of restaurants and their ratings on Yelp.

Aside from analyzing the success and behavior of businesses, we also decided to explore the relationship between the text in reviews and the number of stars it received. In our project, we used a deep dataset of Round 7 of the Yelp Academic Dataset challenge, available at: https://www.yelp.com/dataset_challenge/dataset. The dataset includes five json files of business, check-in, user, review, and tip data, but our research only involved looking at the business, user, and review datasets.

@@

####Stage Two: Data Wrangling
We began by filtering for businesses with more than 300 reviews in the United States. Looking more closely at the business dataset, we noticed a nested data-frame of attributes for each business, so we cleaned and gathered the data so that each row contained one business attribute. Business data set was used to see what makes a business popular and successful; especially by looking at where most-reviewed are clustered and the influence of major business attributes.

Review data was cleaned for sentiment analysis using a library called “tidytext”. After reading in the review dataset, we randomly sampled 500,000 observations. With this, we tokenized each word in a review into a single row in the data-frame and removed stop words like “a”, “and”, “the”, etc. since they had no indication of the reviewers’ attitudes. Using built-in lexicons, we assigned each word a valence score and computed the mean score for each review. Additionally, we created per-word summaries, computing the number of business and reviews the word appeared in and the average star rating based on each review. We filtered for words that appeared in more than 10 businesses and 500 reviews to exclude rare words with strange valence scores and ratings. Using these cleaned datasets, we were able to create visualizations and make inferences discussed later.

```{r load, echo = FALSE}
reviews <- read_csv("../clean_data/yelp_reviews.csv")
reviews <- reviews %>% 
  sample_n(500000)
```

Lastly, business data, user data and review data were all used to analyze the users’ behaviors when they are in unfamiliar places. User data was cleaned to filter for active users with a significant amount of reviews. Cleaned review data and business data were then joined together with user data. We created an additional boolean column called ‘most_reviewed’, after computing the most familiar location for a user (one state per user where most of his/her reviews reside). Using this, we cleaned up and created a new data frame with average star ratings of familiar places and unfamiliar places per user. In the next section, we discuss this further along with results.

####Stage Three: Pre-Processing using ‘tidytext’ for Sentiment Analysis
Sentiment analysis, or opinion mining, makes use of natural language processing to analyze pieces of text and determine the attitude of the speaker or writer. To perform sentiment analysis, we researched various text mining libraries. First, _tm_ was used to remove stop words, common English words that would give no information, but the process of splitting text reviews into one word-per-row and removing stop words was too inefficient. After more research, we came across “tidytext”, a library for text mining. Using this library, cleaning and aggregating the review data was extremely easy and efficient. We then performed tidy analysis to strip reviews and tokenize each word into a single row in the data-frame.

@@

##Results and Discussion
####B. Sentiment Analysis
We hypothesized that words with higher valence scores would produce higher Yelp star ratings for businesses. First, using the R library “wordcloud”, we formed two word clouds for the 100 most frequent words in text reviews with one-star and five-star ratings to determine if there was a distinct difference between the words used.

```{r wordcloud, echo = FALSE, message = FALSE}
reviews_text <- read_csv("../clean_data/reviews_text.csv")
one <- reviews_text %>% 
  filter(stars == 1) %>% 
  group_by(word) %>% 
  count() %>% 
  top_n(100) %>% 
  arrange(n)

@@

@@

```{r boxplots, echo = FALSE, message = FALSE, fig.align = "center"}
review_scores <- read_csv("../clean_data/review_scores.csv")
ggplot(review_scores) +
  geom_boxplot(aes(stars, sentiment, group = stars)) +
  labs(x = "Business Star Rating",
       y = "Mean Sentiment Score",
       title = "Mean Sentiment Score vs. Business Star Rating") +
  theme(text = element_text(family = "Times New Roman"))
```

@@

```{r afinn, echo = F, message = F, fig.align = "center", fig.height = 6}
summaries <- read_csv("../clean_data/summaries.csv")
summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word, family = "Times New Roman"), 
            check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10(limits = c(200, 100000), 
                breaks = c(1000, 10000), labels = c(1000, 10000)) +
  geom_hline(yintercept = mean(reviews$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", 
                        midpoint = 0, mid = "grey") +
  labs(x = "# of reviews",
       y = "Average Yelp Stars",
       color = "Valence Score") +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  theme(text = element_text(family = "Times New Roman"))

positive <- summaries %>% 
  filter(score > 0)
negative <- summaries %>% 
  filter(score < 0)
percent_pos <- PercentAccuracy(positive)
percent_neg <- 1 - PercentAccuracy(negative)
print(str_c("Positivity Accuracy: ", percent_pos))
print(str_c("Negativity Accuracy: ", percent_neg))
```

For the most part, our classifier was accurate. 65.02% of positive words sampled lied above the mean restaurant rating, and even better, 86.99% of negative words lied below the mean star rating. Therefore, it seems reasonable to infer that there is indeed a positive correlation between valence score and Yelp business rating.

@@

```{r trend, echo = FALSE, message = F, fig.align = "center"}
summaries %>% 
  arrange(desc(reviews)) %>% 
  ggplot(aes(score, average_star)) +
  geom_point(aes(size = reviews)) +
  geom_text(aes(label = word, family = "Times New Roman"),
            vjust = 1, hjust = 1, check_overlap = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Valence Score") +
  ylab("Average Yelp Stars") +
  labs(title = "Average Yelp Star Rating vs. Valence Score") +
  scale_x_continuous(limits = c(-5,6), breaks = seq(-5, 5, 2)) +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  expand_limits(x = -6) +
  theme(text = element_text(family = "Times New Roman"))
```

@@

```{r bar, echo = F, message = F, fig.align = "center"}
data.frame(
  x = factor(-5:5), 
  y = unlist(-5:5 %>% map(MeanStar))) %>% 
  filter(!is.nan(y)) %>% 
  ggplot(aes(x = x, y = y, fill = x)) +
  geom_bar(stat = "identity") +
  labs(x = "Valence Score",
       y = "Average Star Rating",
       fill = "Valence Score") +
  theme(text = element_text(family = "Times New Roman"))
```

@@

>>>>

Whenever a person hears about a business or think about starting their own, they are always told that a business’s success is all about location, location, location. It’s hard to quantify how much of a business’s success is determined by its location, but we were curious to see if location was linked to higher star ratings on Yelp. Yelp is a service that publishes crowd-sourced reviews about local business online. Additionally, we wanted to see if there was any relationship between certain attributes of restaurants and their ratings on Yelp. Addiontally, we wanted to explore the the behavior of reviewers when they travel to places outside their "home" location.

Aside from analyzing the success and behavior of reviewers and businesses, we also decided to explore the relationship between the text in reviews and the number of stars it received. In our project, we used a deep dataset of Round 7 of the Yelp Academic Dataset challenge, available at: [https://www.yelp.com/dataset_challenge/dataset]. The dataset includes five json files of business, check-in, user, review, and tip data, but our research only involved looking at the business, user, and review datasets.

@@

####Stage Two: Data Wrangling
We began by filtering for businesses with more than 300 reviews in the United States. Looking more closely at the business dataset, we noticed a nested data-frame of attributes for each business, so we cleaned and gathered the data so that each row contained one business attribute. Business data set was used to see what makes a business popular and successful; especially by looking at where most-reviewed are clustered and the influence of major business attributes.
```{r load_business, echo = FALSE}
business <- read_csv("../clean_data/yelp_business.csv")
names(business) <- c("business_id", "open", "categories", "count", "state", "stars", "noise", "attire", "Take-out", "Takes Reservation", "Offers Delivery", "Outdoor Seating", "Accepts Credit Card", "Happy Hour")
successful_business <- business %>% 
  select(-open) %>% 
  filter(count > 300) %>% 
  group_by(state) %>% 
  summarise(num = n())
```
Review data was cleaned for sentiment analysis using a library called “tidytext”. After reading in the review dataset, we randomly sampled 500,000 observations. With this, we tokenized each word in a text review into a single row in the data-frame and removed stop words like “a”, “and”, “the”, etc. since they had no indication of the reviewers’ attitudes. Using built-in lexicons, we assigned each word a valence score and computed the mean score for each review. Additionally, we created per-word summaries, computing the number of business and reviews the word appeared in and the average star rating based on each review. We filtered for words that appeared in more than 10 businesses and 500 reviews to exclude rare words with strange valence scores and ratings. Using these cleaned datasets, we were able to create visualizations and make inferences discussed later.

```{r load_reviews, echo = FALSE}
reviews <- read_csv("../clean_data/yelp_reviews.csv")
reviews_sample <- reviews %>% 
  sample_n(500000)
```

Lastly, business data, user data and review data were all used to analyze the users’ behaviors when they are in unfamiliar locations User data was cleaned to filter for active users with a significant amount of reviews. Cleaned review data and business data were then joined together with user data. We created an additional boolean column called ‘most_revieSwed’, after computing a user's "home" (state where they reviewed the most). Using this, we cleaned up and created a new data frame with average star ratings of familiar places and unfamiliar places per user. In the next section, we discuss this further along with results. 
```{r load_users, echo = FALSE}
users <- read_csv("../clean_data/yelp_users.csv")
reviews <- read_csv("../clean_data/yelp_reviews.csv")
```
####Stage Three: Pre-Processing using ‘tidytext’ for Sentiment Analysis
Sentiment analysis, or opinion mining, makes use of natural language processing to analyze pieces of text and determine the attitude of the speaker or writer. To perform sentiment analysis, we researched various text mining libraries. First, "__tm__" was used to remove stop words, common English words that would give no information, but the process of splitting text reviews into one word-per-row and removing stop words was too inefficient. After more research, we came across “__tidytext__”, a library for text mining. Using this library, cleaning and aggregating the review data was extremely easy and efficient. We then performed tidy analysis to strip reviews and tokenize each word into a single row in the data-frame.

@@

##Results and Discussion

####A. Location Mining and Business Attributes

Our first hypothesis was that successful businesses would be clustered at certain points. We first tried to plot this information with points on a map using longitude and latitude of the businesses.  Points with different sizes stacked up on each other, so the size and color of the points did not tell much information. Even though 900+ points were plotted, there were really only six primary clusters of points so there was no point in having so many small individual points.

Instead, we ended up with a bar graph, which clearly shows which state has the most number of businesses with over 300 reviews. There were 6 distinct regions in total. Interestingly, the graph shows that the majority of the businesses is located in two states: about 62% of the filtered businesses were located in Nevada and another 31% was in Arizona. 
```{r business, echo = FALSE, fig.align = "center"}
ggplot(successful_business, aes(x = reorder(state, -num), 
                                y= num)) + 
  geom_bar(stat = "identity", 
           position = "dodge") +
  xlab("State") + 
  ylab("Number of Businesses") +
  theme(text = element_text(family = "Times New Roman"))
```

\begin{center}
\textbf{Figure 1:} Number of businesses with more than 300 reviews
\end{center}

We also hypothesized that there were certain business attributes that would help businesses succeed in terms of number of reviews and ratings. We first tried to include two representative plots, ‘Noise Level’ and ‘Attire Type’. However, subsetting too much of the data set made us lose valuable information about the dataset. Choosing only two random attributes and analyzing was too narrow an exploration of the data.

As such, six major attributes were chosen. There were slight differences in star ratings based on some categories. For example, businesses that accept credit cards seemed to have higher ratings compared to those that don't. Interestingly, restaurants that do not offer delivery showed slightly higher star ratings than those who do. There were no major findings from the plots, however. Except for slight differences in average ratings of certain attributes, there was no significant difference evident. The user review ratings were evenly spread out even if the specific attribute existed or not. Our findings pointed out that business attributes do not have huge influences over user ratings.
```{r attributes, echo = FALSE, fig.align = "center"}
business_attributes <- business %>% 
  filter(count > 300) %>% 
  select(-open, -count, -categories, -noise, -attire) %>% 
  gather(key = Attribute, value = Val, -business_id, -state, -stars) %>% 
  na.omit

ggplot(business_attributes, 
  aes(x = factor(Val, levels = c("TRUE","FALSE")), y = stars, fill = state)) + 
  facet_wrap(~Attribute) + 
  geom_bar(stat = "identity", position = "dodge") +
  xlab("") + 
  ylab("Star Rating") +
  scale_fill_discrete(name = "State") +
  theme(text = element_text(family = "Times New Roman"))
```

\begin{center}
\textbf{Figure 2:} User ratings on business attributes
\end{center}

####B. Sentiment Analysis
We hypothesized that words with higher valence scores would produce higher Yelp star ratings for businesses. First, using the R library “wordcloud”, we formed two word clouds for the 100 most frequent words in text reviews with one-star and five-star ratings to determine if there was a distinct difference between the words used.

```{r wordcloud, echo = FALSE, message = FALSE, fig.height = 3.5}
reviews_text <- read_csv("../clean_data/reviews_text.csv")
one <- reviews_text %>% 
  filter(stars == 1) %>% 
  group_by(word) %>% 
  count() %>% 
  top_n(100) %>% 
  arrange(n)

@@

\begin{center}
\textbf{Figure 3:} 100 most common words for one-star and five-star rated businesses
\end{center}

@@

```{r boxplots, echo = FALSE, message = FALSE, fig.align = "center"}
review_scores <- read_csv("../clean_data/review_scores.csv")
ggplot(review_scores) +
  geom_boxplot(aes(stars, sentiment, group = stars)) +
  labs(x = "Business Star Rating",
       y = "Mean Sentiment Score") +
  theme(text = element_text(family = "Times New Roman"))
```
\begin{center}
\textbf{Figure 4:} Mean Sentiment Score vs. Business Star Rating
\end{center}

@@

```{r afinn, echo = F, message = F, fig.align = "center", fig.height = 6}
summaries <- read_csv("../clean_data/summaries.csv")
summaries %>% 
  ggplot(aes(reviews, average_star, color = score)) +
  geom_point() +
  geom_text(aes(label = word, family = "Times New Roman"), 
            check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10(limits = c(200, 100000), 
                breaks = c(1000, 10000), labels = c(1000, 10000)) +
  geom_hline(yintercept = mean(reviews_sample$stars), color = "black", lty = 2) +
  scale_color_gradient2(low = "red", high = "blue", 
                        midpoint = 0, mid = "grey") +
  labs(x = "# of reviews",
       y = "Average Yelp Stars",
       color = "Valence Score") +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  theme(text = element_text(family = "Times New Roman"))

positive <- summaries %>% 
  filter(score > 0)
negative <- summaries %>% 
  filter(score < 0)
percent_pos <- PercentAccuracy(positive)
percent_neg <- 1 - PercentAccuracy(negative)
print(str_c("Positivity Accuracy: ", percent_pos))
print(str_c("Negativity Accuracy: ", percent_neg))
```
\begin{center}
\textbf{Figure 5:} Average Yelp Rating vs. Frequency of Words
\end{center}

For the most part, our classifier was accurate. 65.02% of positive words sampled lied above the mean restaurant rating, and even better, 86.59% of negative words lied below the mean star rating. Therefore, it seems reasonable to infer that there is indeed a positive correlation between valence score and Yelp business rating.

@@

```{r trend, echo = FALSE, message = F, fig.align = "center"}
summaries %>% 
  arrange(desc(reviews)) %>% 
  ggplot(aes(score, average_star)) +
  geom_point(aes(size = reviews)) +
  geom_text(aes(label = word, family = "Times New Roman"),
            vjust = 1, hjust = 1, check_overlap = TRUE) +
  geom_smooth(method = "lm", se = FALSE) +
  xlab("Valence Score") +
  ylab("Average Yelp Stars") +
  scale_x_continuous(limits = c(-5,6), breaks = seq(-5, 5, 2)) +
  scale_y_continuous(limits = c(1, 5), breaks = seq(1, 5, 1)) +
  expand_limits(x = -6) +
  theme(text = element_text(family = "Times New Roman"))
```
\begin{center}
\textbf{Figure 6:} Average Yelp Star Rating vs. Valence Score
\end{center}

@@

```{r bar, echo = F, message = F, fig.align = "center"}
data.frame(
  x = factor(-5:5), 
  y = unlist(-5:5 %>% map(MeanStar))) %>% 
  filter(!is.nan(y)) %>% 
  ggplot(aes(x = x, y = y, fill = x)) +
  geom_bar(stat = "identity") +
  labs(x = "Valence Score",
       y = "Average Star Rating",
       fill = "Valence Score") +
  theme(text = element_text(family = "Times New Roman"))
```
\begin{center}
\textbf{Figure 7:} Bar Graph of Average Yelp Rating per AFINN sentiment score
\end{center}

@@

####C. User Behavior when Travelling

Looking at the user and review datasets combined, we wanted to find out what would affect people's behavior of rating businesses. Since the dataset given by Yelp was academic, this did not include every single review by every user in data. We filtered to focus on active users that have written more than 50 reviews in total and have 10 reviews. This seemed like a reasonable constraint on the data that would remove any outlier behavior and would still capture the overall trend in user behavior when travelling.

We hypothesized that users would give relatively higher star ratings when travelling than when visiting local restaurants that they are familiar with.

First, we faceted the graph after selecting the top 15 users with most reviews. However, since facet focuses rather on individual data than overall trends, it was hard to conceputalize the findings. Individual patterns varied, so to capture the overall trend, we plotted their behaviors into one graph.

Another approach we took was to plot users' ratings in a histogram. From the filtered user dataset, one familiar region was chosen per user. This was based on where the user wrote the most reviews. Two average star ratings were calculated for a user: one for reviews in familiar state and another from reviews when travelling to unfamiliar places. The best visualization for such a graph was using a density plot, shown below.

```{r density, echo = FALSE, fig.align = "center"}
selected_users <- users %>% 
  filter(review_count > 50) %>% 
  select(name, user_id, review_count)

user_reviews <- reviews %>% 
  select(user_id, review_id, stars, business_id)

business_locations <- business %>% 
  select(business_id, state)

user_reviews <- left_join(selected_users, user_reviews) %>% 
  left_join(business_locations) %>% 
  na.omit

familiar_location <- group_by(user_reviews, user_id, state) %>% 
  tally()

location_reviews <- aggregate(n ~ user_id, familiar_location, max) %>%
  left_join(familiar_location) %>% 
  mutate(most_reviewed = TRUE) %>% 
  filter(n > 10) %>% 
  select(-n)

user_reviews <- left_join(user_reviews, location_reviews) 
user_reviews[is.na(user_reviews)] <- FALSE

# average stars for unfamiliar/familiar area for a user
avg_star_reviews <- user_reviews %>% 
  group_by(user_id, most_reviewed) %>% 
  summarise(avg_star = mean(stars))

ggplot(avg_star_reviews, aes(x = avg_star, 
                             fill = most_reviewed)) + 
  geom_density(alpha = 0.5,
               color = NA) +
  labs(title = "Reviewers' Behavior Changes When In Unfamilar Locations") +
  xlab("Average Star Ratings") + 
  ylab("Density") +
  scale_fill_manual(name ="Regions",
                    values = c("blue", "red"),
                    labels = c("Familiar", "Unfamiliar"))
```
\begin{center}
\textbf{Figure 8: "Reviewers' Behavior Changes When In Unfamilar Locations"}
\end{center}

The density curve for unfamiliar region is smooth and the majority of users tend to give star ratings in between 3 to 4.5 out of 5. Number of people giving 1 or 2 stars while traveling is barely seen. On the other hand, the ratings in familiar places were rather widely distributed. There was a significant portion of users giving star ratings from 1 through 5, even though the peak of the curve is around 4.

The mean rating was 3.718 for unfamiliar and 3.762 for familiar places. However, the standard deviation for the familiar locations plot was 0.41, and unfamiliar one was 1.37 so star rating, explained by the variablity between users who travelled.

##Conclusion
In conclusion, we investigated several questions related to the Yelp datasets, and our results were promising.

Using natural language processing and text mining, we confirmed our original hypothesis that a positive correlation exists between AFINN sentiment score and star rating. 86.59% of words with negative scores appeared in reviews that were below the mean rating of all reviews from our sample.

##Future Work
We would use cross-validation to create a training set and test set for our sample of reviews from the dataset. We can create a classifier using sentiment analysis and test our trained classifier against the test set to predict business ratings based on the text of the reviews. Additionally, we could look at possible correlations between restaurant reviews and ratings and tips, looking at the tips data provided with the dataset.
